### 2016-12-07 Paper_v1.tex
Dec6

Gene knockouts in the electron transport chain significantly reduced the performance of the CFPS reaction (Fig.~\ref{fig:oxphos_ko}).
A key finding of both the CAT and overall system state sensitivity analysis was the importance of oxidative phosphorylation.
To investigate this further, we knocked out key oxidative phosphorylation reactions in the ensemble of kinetic models to examine the effect on glucose uptake and CAT production.
%A single \textit{cyd} knockout reduced the CAT carbon yield from 7.9\% to 2.6\% (Table~\ref{tbl:yield_breakdown}).
A single \textit{cyd} knockout reduced the CAT carbon yield from 2.7\% to 0.9\% (Table~\ref{tbl:yield_breakdown}).
In addition, the glucose uptake rate was reduced compared to that of the control (no knockouts).
%On the other hand, a \textit{nuo} knockout showed a less dramatic decrease in yield,
%reducing the CAT carbon yield to 6.9\%; however, the glucose uptake rate remained similar to that of the control.
On the other hand, a \textit{nuo} knockout showed a less dramatic decrease in yield,
reducing the CAT carbon yield to 2.3\%; however, the glucose uptake rate remained similar to that of the control.
%Knocking out \textit{app} increased CAT yield to 8.3\%, but this increase was not statistically significantly different from that of the control.
Knocking out \textit{app} did not change the CAT yield (it remained at 2.7\%).
%Lastly, knocking out all three reactions reduced the CAT yield to 3.0\%, but this was not statistically significantly different from the \textit{cyd} knockout alone.
Lastly, knocking out all three reactions reduced the CAT yield to 0.7\%, similar to knocking out the \textit{cyd} alone.
Thus, the model suggested the key step in oxidative phosphorylation was catalyzed by the gene product of \textit{cyd}.
However, while disruption of \textit{cyd} significantly reduced the CAT carbon yield, it did not eliminate the ability of CFPS reaction to produce CAT.
This suggested there was a mixture of energy sources supporting CAT production, with the most significant being oxidative phosphorylation.
%The \textit{cyd} enzyme is responsible for transferring electrons with oxygen as an electron acceptor allowing the release of energy to form ATP.
%The sensitivity analysis showed oxidative phosphorylation was significant for CAT production as well as the system state.
A similar distribution of the carbon contribution to CAT yield was seen across the best-fit set and all knockouts.
In all cases, glutamate was responsible for about two-thirds of carbon consumption toward CAT.
This was due to its role in generating other amino acids and as a possible energy source for metabolism.
The much greater consumption of this one amino acid was made possible by its much larger initial condition, as it was present in the cell media in the form of magnesium glutamate, ammonium glutamate, and potassium glutamate [CITE? THESIS NOT CURRENTLY IN BIBLIOGRAPHY].
Glucose made the next largest contribution, between 20\% and 30\%, dwarfing all amino acids other than glutamate.
This makes sense, as glucose powers the entire metabolism during the first hour, including energy species synthesis and amino acid synthesis.

@@

@@

\begin{table}
\centering
    \caption{CAT carbon yield breakdown for best-fit set, knockouts, and experimental data. Carbon produced as CAT, carbon consumed as glucose and each amino acid, sum of consumed species (with and without arginine and glutamate), and yield  (with and without arginine and glutamate).}
    \renewcommand{\arraystretch}{1.3}
    \begin{tabular}{lrrrrrr} \toprule
        Carbon Produced (mM) & Best-fit & $\Delta$cyd & $\Delta$nuo & $\Delta$app & $\Delta$cyd $\Delta$nuo $\Delta$app & Data \\ \hline
        CAT & 20.9 & 6.5 & 18.1 & 21.4 & 5.1 & 21.6 \\ \midrule
        Carbon Consumed (mM) \\ \hline
        GLC & 215.4 & 215.4 & 215.4 & 215.4 & 159.8 & 215.4 \\ \hline
        ALA & 0.0 & 0.0 & 1.7 & 0.0 & 0.0 & 0.0 \\ \hline
        ARG & 10.2 & 0.9 & 1.1 & 9.9 & 1.3 & - \\ \hline
        ASN & 6.2 & 6.3 & 6.2 & 6.2 & 6.3 & 6.3 \\ \hline
        ASP & 7.5 & 0.0 & 3.9 & 7.5 & 0.0 & 9.6 \\ \hline
        CYS & 3.0 & 2.9 & 3.0 & 3.1 & 2.9 & 3.7 \\ \hline
        GLN & 0.0 & 1.8 & 0.0 & 0.0 & 2.7 & 0.0 \\ \hline
        GLU & 492.6 & 505.1 & 528.2 & 505.6 & 501.8 & - \\ \hline
        GLY & 3.1 & 1.1 & 2.6 & 3.1 & 0.9 & 1.5 \\ \hline
        HIS & 0.2 & 0.4 & 1.1 & 0.2 & 0.3 & 0.0 \\ \hline
        ILE & 1.0 & 0.3 & 0.8 & 1.0 & 0.2 & 1.7 \\ \hline
        LEU & 1.4 & 0.4 & 1.2 & 1.4 & 0.3 & 2 \\ \hline
        LYS & 10.7 & 13.2 & 13.1 & 10.7 & 13.2 & 8.3 \\ \hline
        MET & 0.8 & 0.2 & 0.7 & 0.8 & 0.2 & 2.9 \\ \hline
        PHE & 3.2 & 1.0 & 2.8 & 3.3 & 0.8 & 1.6 \\ \hline
        PRO & 2.4 & 0.2 & 0.7 & 2.4 & 0.2 & 1.9 \\ \hline
        SER & 2.5 & 2.1 & 2.4 & 2.5 & 2.1 & 1.8 \\ \hline
        THR & 3.4 & 2.9 & 3.3 & 3.4 & 2.8 & 2.8 \\ \hline
        TRP & 1.0 & 0.3 & 0.8 & 1.0 & 0.2 & 1.2 \\ \hline
        TYR & 1.1 & 0.4 & 1.1 & 1.1 & 0.4 & 0.6 \\ \hline
        VAL & 1.4 & 0.4 & 1.2 & 1.5 & 0.4 & 2.4 \\ \midrule
        Sum & 767.1 & 755.3 & 791.3 & 780.1 & 696.8 & - \\ \hline
        Sum w/o ARG,GLU & 264.3 & 249.3 & 262.0 & 264.6 & 193.7 & 263.7 \\ \midrule
        Yield & 2.7\% & 0.9\% & 2.3\% & 2.7\% & 0.7\% & - \\ \hline
        Yield w/o ARG, GLU & 7.9\% & 2.6\% & 6.9\% & 8.1\% & 2.7\% & 8.2\% \\ \bottomrule
    \end{tabular}
\label{tbl:yield_breakdown}
\end{table}

\begin{table}
\centering
    \caption{Amino acid consumption toward CAT as a percentage of total amino acid consumption, for the best-fit set. Total consumption defined as decrease in amino acid concentration over 3 hours, such that for amino acids with synthesis active, percentage may be greater than 100\%. Alanine and glutamine did not decrease over the timecourse.}
    \renewcommand{\arraystretch}{1.3}
    \begin{tabular}{lrrr} \toprule
        Species & Total Consumption & Consumption Toward CAT & Percentage \\ \hline
%        ALA  & -3.852  & 0.267  & -6.9\% \\ \hline
        ALA  & - & 0.27 & - \\ \hline
%        ARG  & 1.697  & 0.178  & 10.5\% \\ \hline
%        ASN  & 1.547  & 0.089  & 5.8\% \\ \hline
%        ASP  & 1.875  & 0.214  & 11.4\% \\ \hline
%        CYS  & 1.016  & 0.089  & 8.8\% \\ \hline
        ARG  & 1.70  & 0.18  & 11\% \\ \hline
        ASN  & 1.55  & 0.09  & 6\% \\ \hline
        ASP  & 1.87  & 0.21  & 11\% \\ \hline
        CYS  & 1.02  & 0.09  & 9\% \\ \hline
%        GLN  & -2.278  & 0.214  & -9.4\% \\ \hline
        GLN  & - & 0.21 & - \\ \hline
%        GLU  & 98.526  & 0.232  & 0.2\% \\ \hline
%        GLY  & 1.545  & 0.178  & 11.5\% \\ \hline
%        HIS  & 0.036  & 0.214  & 587.9\% \\ \hline
%        ILE  & 0.16  & 0.16  & 100\% \\ \hline
%        LEU  & 0.231  & 0.232  & 100.2\% \\ \hline
%        LYS  & 1.785  & 0.214  & 12.0\% \\ \hline
%        MET  & 0.16  & 0.16  & 100\% \\ \hline
%        PHE  & 0.356  & 0.356  & 100\% \\ \hline
%        PRO  & 0.481  & 0.125  & 26\% \\ \hline
%        SER  & 0.83  & 0.178  & 21.5\% \\ \hline
%        THR  & 0.857  & 0.232  & 27\% \\ \hline
%        TRP  & 0.089  & 0.089  & 100\% \\ \hline
%        TYR  & 0.12  & 0.196  & 164\% \\ \hline
%        VAL  & 0.285  & 0.285  & 100\% \\ \hline
        GLU  & 98.53  & 0.23  & 0.2\% \\ \hline
        GLY  & 1.55  & 0.18  & 12\% \\ \hline
        HIS  & 0.04  & 0.21  & 588\% \\ \hline
        ILE  & 0.16  & 0.16  & 100\% \\ \hline
        LEU  & 0.23  & 0.23  & 100\% \\ \hline
        LYS  & 1.79  & 0.21  & 12\% \\ \hline
        MET  & 0.16  & 0.16  & 100\% \\ \hline
        PHE  & 0.36  & 0.36  & 100\% \\ \hline
        PRO  & 0.48  & 0.12  & 26\% \\ \hline
        SER  & 0.83  & 0.18  & 21\% \\ \hline
        THR  & 0.86  & 0.23  & 27\% \\ \hline
        TRP  & 0.09  & 0.09  & 100\% \\ \hline
        TYR  & 0.12  & 0.20  & 164\% \\ \hline
        VAL  & 0.29  & 0.29  & 100\% \\ \bottomrule
    \end{tabular}
\label{tbl:AA_breakdown}
\end{table}

>>>>

Gene knockouts in the electron transport chain significantly reduced the performance of the CFPS reaction (Fig.~\ref{fig:oxphos_ko}).
A key finding of both the CAT and overall system state sensitivity analysis was the importance of oxidative phosphorylation.
To investigate this further, we knocked out key oxidative phosphorylation reactions in the ensemble of kinetic models to examine the effect on glucose uptake and CAT production.
%A single \textit{cyd} knockout reduced the CAT carbon yield from 7.9\% to 2.6\% (Table~\ref{tbl:yield_breakdown}).
A single \textit{cyd} knockout reduced the CAT carbon yield from 2.7\% to 0.9\% (Table~\ref{tbl:yield_breakdown}).
In addition, the glucose uptake rate was reduced compared to that of the control (no knockouts).
%On the other hand, a \textit{nuo} knockout showed a less dramatic decrease in yield,
%reducing the CAT carbon yield to 6.9\%; however, the glucose uptake rate remained similar to that of the control.
On the other hand, a \textit{nuo} knockout showed a less dramatic decrease in yield,
reducing the CAT carbon yield to 2.3\%; however, the glucose uptake rate remained similar to that of the control.
%Knocking out \textit{app} increased CAT yield to 8.3\%, but this increase was not statistically significantly different from that of the control.
Knocking out \textit{app} did not change the CAT yield (it remained at 2.7\%).
%Lastly, knocking out all three reactions reduced the CAT yield to 3.0\%, but this was not statistically significantly different from the \textit{cyd} knockout alone.
Lastly, knocking out all three reactions reduced the CAT yield to 0.7\%, similar to knocking out the \textit{cyd} alone.
Thus, the model suggested the key step in oxidative phosphorylation was catalyzed by the gene product of \textit{cyd}.
However, while disruption of \textit{cyd} significantly reduced the CAT carbon yield, it did not eliminate the ability of CFPS reaction to produce CAT.
This suggested there was a mixture of energy sources supporting CAT production, with the most significant being oxidative phosphorylation.
%The \textit{cyd} enzyme is responsible for transferring electrons with oxygen as an electron acceptor allowing the release of energy to form ATP.
%The sensitivity analysis showed oxidative phosphorylation was significant for CAT production as well as the system state.
A similar distribution of the carbon contribution to CAT yield was seen across the best-fit set and all knockouts.
In all cases, glutamate was responsible for about two-thirds of carbon consumption toward CAT.
This was due to its role in generating other amino acids and as a possible energy source for metabolism.
The much greater consumption of this one amino acid was made possible by its much larger initial condition, as it was present in the cell media in the form of magnesium glutamate, ammonium glutamate, and potassium glutamate [CITE? THESIS NOT CURRENTLY IN BIBLIOGRAPHY].
Glucose made the next largest contribution, between 20\% and 30\%, dwarfing all amino acids other than glutamate.
This makes sense, as glucose powers the entire metabolism during the first hour, including energy species synthesis and amino acid synthesis.
The knockouts in oxidative phosphorylation also reduced the energy efficiency of CAT production  (Fig.~\ref{fig:oxphos_ko}).
The best-fit set, $\Delta$app, and experimental dataset had efficiencies of 1.8-1.9\%, $\Delta$nuo had an efficiency of 1.6\%, and $\Delta$cyd and $\Delta$cyd$\Delta$nuo$\Delta$app had efficiencies of 0.6\%.
This was in line with the effect of knockouts on yield, which in turn was more or less proportional to the total amount of CAT produced in each case.
Taken together, these results show that oxidative phosphorylation, particularly \textit{cyd}, is important to efficient CAT production.

@@

Energy efficiency was also calculated for best-fit set, knockouts, and experimental data, formulated as:
\begin{equation}\label{eqn:energy-efficiency-definition}
	\rm Efficiency=\frac{\Delta\texttt{CAT}\cdot \left(2*\rm {ATP}_{CAT}+\rm {GTP}_{CAT}\right)}{\Delta\texttt{GLC}\cdot \rm {ATP}_{GLC}}
\end{equation}
where $\rm Efficiency$ denotes the energy efficiency of CAT production, $\rm {ATP}_{CAT}$ denotes the stoichiometric coefficient of ATP in CAT synthesis (multiplied by 2 because AMP, rather than ADP, is a product of CAT synthesis), $\rm {GTP}_{CAT}$ denotes the stoichiometric coefficient of GTP in CAT synthesis, and $\rm {ATP}_{GLC}$ denotes the number of ATP molecules ideally produced from one molecule of glucose.
$\rm {ATP}_{CAT}$ = 219, $\rm {GTP}_{CAT}$ = 438, $\rm {ATP}_{GLC}$ = 24.

@@

\begin{table}
\centering
    \caption{CAT carbon yield breakdown for best-fit set, knockouts, and experimental data. Carbon produced as CAT, carbon consumed as glucose and each amino acid, sum of consumed species (with and without arginine and glutamate), and yield  (with and without arginine and glutamate). Energy efficiency for best-fit set, knockouts, and experimental data.}
    \renewcommand{\arraystretch}{1.25}
    \begin{tabular}{lrrrrrr} \toprule
        Carbon Produced (mM) & Best-fit & $\Delta$cyd & $\Delta$nuo & $\Delta$app & $\Delta$cyd $\Delta$nuo $\Delta$app & Data \\ \hline
        CAT & 20.9 & 6.5 & 18.1 & 21.4 & 5.1 & 21.6 \\ \midrule
        Carbon Consumed (mM) \\ \hline
        GLC & 215.4 & 215.4 & 215.4 & 215.4 & 159.8 & 215.4 \\ \hline
        ALA & 0.0 & 0.0 & 1.7 & 0.0 & 0.0 & 0.0 \\ \hline
        ARG & 10.2 & 0.9 & 1.1 & 9.9 & 1.3 & - \\ \hline
        ASN & 6.2 & 6.3 & 6.2 & 6.2 & 6.3 & 6.3 \\ \hline
        ASP & 7.5 & 0.0 & 3.9 & 7.5 & 0.0 & 9.6 \\ \hline
        CYS & 3.0 & 2.9 & 3.0 & 3.1 & 2.9 & 3.7 \\ \hline
        GLN & 0.0 & 1.8 & 0.0 & 0.0 & 2.7 & 0.0 \\ \hline
        GLU & 492.6 & 505.1 & 528.2 & 505.6 & 501.8 & - \\ \hline
        GLY & 3.1 & 1.1 & 2.6 & 3.1 & 0.9 & 1.5 \\ \hline
        HIS & 0.2 & 0.4 & 1.1 & 0.2 & 0.3 & 0.0 \\ \hline
        ILE & 1.0 & 0.3 & 0.8 & 1.0 & 0.2 & 1.7 \\ \hline
        LEU & 1.4 & 0.4 & 1.2 & 1.4 & 0.3 & 2 \\ \hline
        LYS & 10.7 & 13.2 & 13.1 & 10.7 & 13.2 & 8.3 \\ \hline
        MET & 0.8 & 0.2 & 0.7 & 0.8 & 0.2 & 2.9 \\ \hline
        PHE & 3.2 & 1.0 & 2.8 & 3.3 & 0.8 & 1.6 \\ \hline
        PRO & 2.4 & 0.2 & 0.7 & 2.4 & 0.2 & 1.9 \\ \hline
        SER & 2.5 & 2.1 & 2.4 & 2.5 & 2.1 & 1.8 \\ \hline
        THR & 3.4 & 2.9 & 3.3 & 3.4 & 2.8 & 2.8 \\ \hline
        TRP & 1.0 & 0.3 & 0.8 & 1.0 & 0.2 & 1.2 \\ \hline
        TYR & 1.1 & 0.4 & 1.1 & 1.1 & 0.4 & 0.6 \\ \hline
        VAL & 1.4 & 0.4 & 1.2 & 1.5 & 0.4 & 2.4 \\ \midrule
        Sum & 767.1 & 755.3 & 791.3 & 780.1 & 696.8 & - \\ \hline
        Sum w/o ARG,GLU & 264.3 & 249.3 & 262.0 & 264.6 & 193.7 & 263.7 \\ \midrule
        Yield & 2.7\% & 0.9\% & 2.3\% & 2.7\% & 0.7\% & - \\ \hline
        Yield w/o ARG, GLU & 7.9\% & 2.6\% & 6.9\% & 8.1\% & 2.7\% & 8.2\% \\ \midrule
        Energy Efficiency & 1.8\% & 0.6\% & 1.6\% & 1.8\% & 0.6\% & 1.9\% \\ \bottomrule
%        Efficiencies based on ATP number of 24: 1.811977 & 0.558704 & 1.565814 & 1.849871 & 0.600006 & 1.873067
    \end{tabular}
\label{tbl:yield_breakdown}
\end{table}

\begin{table}
\centering
    \caption{Amino acid consumption toward CAT as a percentage of total amino acid consumption, for the best-fit set. Total consumption defined as decrease in amino acid concentration over 3 hours, such that for amino acids with synthesis active, percentage may be greater than 100\%. Alanine and glutamine did not decrease over the timecourse.}
    \renewcommand{\arraystretch}{1.25}
    \begin{tabular}{lrrr} \toprule
        Species & Total Consumption & Consumption Toward CAT & Percentage \\ \hline
%        ALA  & -3.852  & 0.267  & -6.9\% \\ \hline
        ALA  & - & 0.27 & - \\ \hline
%        ARG  & 1.697  & 0.178  & 10.5\% \\ \hline
%        ASN  & 1.547  & 0.089  & 5.8\% \\ \hline
%        ASP  & 1.875  & 0.214  & 11.4\% \\ \hline
%        CYS  & 1.016  & 0.089  & 8.8\% \\ \hline
        ARG  & 1.70  & 0.18  & 11\% \\ \hline
        ASN  & 1.55  & 0.09  & 6\% \\ \hline
        ASP  & 1.87  & 0.21  & 11\% \\ \hline
        CYS  & 1.02  & 0.09  & 9\% \\ \hline
%        GLN  & -2.278  & 0.214  & -9.4\% \\ \hline
        GLN  & - & 0.21 & - \\ \hline
%        GLU  & 98.526  & 0.232  & 0.2\% \\ \hline
%        GLY  & 1.545  & 0.178  & 11.5\% \\ \hline
%        HIS  & 0.036  & 0.214  & 587.9\% \\ \hline
%        ILE  & 0.16  & 0.16  & 100\% \\ \hline
%        LEU  & 0.231  & 0.232  & 100.2\% \\ \hline
%        LYS  & 1.785  & 0.214  & 12.0\% \\ \hline
%        MET  & 0.16  & 0.16  & 100\% \\ \hline
%        PHE  & 0.356  & 0.356  & 100\% \\ \hline
%        PRO  & 0.481  & 0.125  & 26\% \\ \hline
%        SER  & 0.83  & 0.178  & 21.5\% \\ \hline
%        THR  & 0.857  & 0.232  & 27\% \\ \hline
%        TRP  & 0.089  & 0.089  & 100\% \\ \hline
%        TYR  & 0.12  & 0.196  & 164\% \\ \hline
%        VAL  & 0.285  & 0.285  & 100\% \\ \hline
        GLU  & 98.53  & 0.23  & 0.2\% \\ \hline
        GLY  & 1.55  & 0.18  & 12\% \\ \hline
        HIS  & 0.04  & 0.21  & 588\% \\ \hline
        ILE  & 0.16  & 0.16  & 100\% \\ \hline
        LEU  & 0.23  & 0.23  & 100\% \\ \hline
        LYS  & 1.79  & 0.21  & 12\% \\ \hline
        MET  & 0.16  & 0.16  & 100\% \\ \hline
        PHE  & 0.36  & 0.36  & 100\% \\ \hline
        PRO  & 0.48  & 0.12  & 26\% \\ \hline
        SER  & 0.83  & 0.18  & 21\% \\ \hline
        THR  & 0.86  & 0.23  & 27\% \\ \hline
        TRP  & 0.09  & 0.09  & 100\% \\ \hline
        TYR  & 0.12  & 0.20  & 164\% \\ \hline
        VAL  & 0.29  & 0.29  & 100\% \\ \bottomrule
    \end{tabular}
\label{tbl:AA_breakdown}
\end{table}
